---
title: "README"
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(JDJcensus)
library(tidyverse)
```

This package includes tools to make accessing census data faster. It wraps commands from the excellent [**tidycensus**](https://walkerke.github.io/tidycensus/index.html) package, which pulls from the Census API. If you haven't already, you'll need to install a census API key. [**Sign up for an API key here**](https://api.census.gov/data/key_signup.html). Then [**follow these instructions**](https://walkerke.github.io/tidycensus/reference/census_api_key.html) to install it in your .Renviron file for repeated use.

`JDJcensus` builds on `tidycensus` by providing a built-in dataframe with all variable names from different data sources combined. It also returns census data with the data source, year, and text label appended.

## Getting started

Install the package this way.

`devtools::install_github("jdjohn215/JDJcensus")`

`JDJcensus` includes a built in file showing all the available variables from 2010-2018 American Community Surveys (ACS) and the 1990, 2000, 2010 SF1 decennial census products. You can access it by running `View(AllVariables)`. It looks like this

```{r, echo=FALSE}
reactable::reactable(AllVariables, filterable = TRUE, searchable = TRUE, defaultPageSize = 5,
                     striped = TRUE)
```

## Downloading data

You can identify the table you want to download either with "table" code or the "concept" label. For example, the `SEX BY AGE` concept is table code `B01001`.

Use either the table code or concept label in the `table_download` function.

This example downloads the selected table for census tracts in Milwaukee County, Wisconsin from the 2018 5YR ACS.
```{r}
table_download("HISPANIC OR LATINO ORIGIN BY RACE",
               geography = "tract", source = "acs", year = 2018, 
               state = "WI", county = "MILWAUKEE")
```

This example downloads the same table for each state in the 2010 decennial census.
```{r}
table_download("HISPANIC OR LATINO ORIGIN BY RACE",
               geography = "state", source = "sf1", year = 2010)
```

Commonly, I find myself downloading data for the same area over and over again. Here's an example of a simple helper function that can speed up that process.
```{r}
quick_acs <- function(table){
  table_download(table = table, geography = "tract", source = "acs", year = 2018,
                 state = "WI", county = "MILWAUKEE")
}

quick_acs("MEDIAN AGE BY SEX")
```
